---
title: "Income Inequality and Environmental Performance"
author: "Raju Adhikari and Mario Rodriguez"
date: "2/12/2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(WDI)
library(countrycode)
library(rio)
library(stargazer)
library(ggplot2)
library(tidyr)
library(plotly)
library(plm)
library(googleVis)

possible_dir <- c('C:/RajuPC/CollaborativeSSDA/Assignments/Website', '/Users/mariorodriguez/Desktop/Website')
repmis::set_valid_wd(possible_dir)

Combined <- read.csv('EPIGINI.csv')

Combined2012 <- read.csv('EPIGINI2012.csv')

names(Combined2012)

Combined2012 <- Combined2012[, c('GINI', 'GDPPerCapPPP', 'CO2EmPerCap', 'EPI.2012', 'iso2c')]

Combined2010 <- read.csv('EPIGINI2010.csv')

GINImap <- gvisGeoChart(Combined2010, locationvar = 'iso2c',
                      colorvar = 'GINI',
                      options = list(
                          colors = "['lightblue', 'blue']"
                          ))

EPImap <- gvisGeoChart(Combined2010, locationvar = 'iso2c',
                      colorvar = 'EPI.2010',
                      options = list(
                          colors = "['red', 'yellow']"
                          ))

GINImap2012 <- gvisGeoChart(Combined2012, locationvar = 'iso2c',
                      colorvar = 'GINI',
                      options = list(
                          colors = "['lightblue', 'blue']"
                          ))

EPImap2012 <- gvisGeoChart(Combined2012, locationvar = 'iso2c',
                      colorvar = 'EPI.2012',
                      options = list(
                          colors = "['red', 'yellow']"
                          ))

labels <- c('Gini Coefficient', 'GDP Per Capita (PPP)', '(Intecept)')
m1 <- lm(EPI.2012 ~ GINI, Combined2012)
m2 <- lm(EPI.2012 ~ GINI + GDPPerCapPPP, Combined2012)
m3 <- lm(EPIValue ~ GiniCoeff + GDPperCapPPP, Combined)
m4 <- plm(EPIValue ~ GiniCoeff + GDPperCapPPP, data=Combined, index= c('country', 'year'), model = 'within')
m5 <- plm(EPIValue ~ GiniCoeff + GDPperCapPPP, data=Combined, index= c('country', 'year'), model = 'random')
```

## Table of contents

- Research Question
- Theoretical Basis
- Operationalization
- Plots, Maps and other Descriptive Statistics
- Regression and Other Inferential Statistics
- Conclusion

## Research question

Unrefined Question:

Does income inequality have an effect on the environment?

Refined Question:

Does income inequality of a country affect its Environment Performance Index? If so, How?

## Theoretical premise

- Carbon emission and equal countries
- Economic Development and unequal countries
- Environment Performance Index

## Operationalization

- Explanatory Variable: 
    - Gini coefficient (measure of income inequality)
    - Source: World Bank's API
- Dependent Variable: 
    - EPI (measure of environment performance)
    - Source: Yale Center for Environmental Law and Policy
- Data Cleaning:
    - Excel File, Yearly data separated by sheets 
    - Columns included items used to calculate EPI
    - Merge the data-sets
- Control Variables:
    - GDP per capita (PPP)

## GINI Map 2010 {.flexbox .vcenter}

```{r descriptive 2, results='asis'}
print(GINImap, tag = 'chart')
```

## EPI Map 2010 {.flexbox .vcenter}

```{r descriptive 3, results='asis'}
print(EPImap, tag = 'chart')
```
## GINI Map 2012 {.flexbox .vcenter}

```{r descriptive 4, results='asis'}
print(GINImap2012, tag = 'chart')
```

## EPI Map 2012 {.flexbox .vcenter}

```{r descriptive 5, results='asis'}
print(EPImap2012, tag = 'chart')
```
## Descriptive statistics  {.flexbox .vcenter}

```{r descriptive, echo = FALSE, results='asis', message = FALSE, warning = FALSE}
stargazer(Combined2012, type = "html", title="Descriptive statistics", digits=1)
```

## Relationship EPI vs. GINI 2012  {.flexbox .vcenter}

```{r okit, echo = FALSE, results='asis', message = FALSE, warning = FALSE}
plot_ly(Combined2012, x = ~GINI, y = ~EPI.2012,
        mode = 'markers')
```

## Inferential statistics {.flexbox .vcenter .smaller}

```{r regression1, results='asis'}
stargazer::stargazer(m1, m2, type='html', covariate.labels = labels, 
          title = 'Effect of Income Inequality on EPI', font.size = 'tiny')
```

## Inferential statistics(contd.) {.flexbox .vcenter .smaller}

```{r regression2, results='asis'}
stargazer::stargazer(m3, m4, m5, type='html', covariate.labels = labels, digits = 2, 
          title = 'Effect of Income Inequality on EPI', font.size = 'normalsize')
```

## Inferential statistics(contd.)

```{r hausman}
phtest(m4, m5)
```

## Conclusion
- Inequality has a negative correlation with EPI
- 1 unit increase in Gini coefficient decreases the country's EPI by 0.1 unit.
- Policies should be directed towards increasing GDP Per Capita (PPP) while making sure that inequality doesn't increase. (Economic Policy to deal with Environment issues)
```{r conclusion}

```
